---
title: "HW4"
author: "Corinna Hong"
date: "February 24, 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center")
```

```{r}

# load packages

library(tidyverse)
library(janitor)
library(paletteer)

```

##1. Visualizing patterns of change for climate change data
```{r}

# read in and wrangle climate change data

climate_change <- read_csv("climate_change.csv") %>% 
  clean_names() %>% 
  mutate(geographic_range = str_to_lower(geographic_range))

```

```{r}

# Visualize changes in density due to development mode

climate_dev_bar <- ggplot(climate_change,
                          aes(x = species,
                              y = change_in_density,
                              fill = development_mode)) +
  geom_bar (stat = "identity") + # identity means height of bars equals actual value
   scale_fill_paletteer_d("ghibli::MononokeMedium", direction = -1) + # direction -1 means start from the right side of the palette
  theme_minimal()

climate_dev_bar



# Visualize changes in density due to geographic range

climate_geo_bar <- ggplot(climate_change,
                          aes(x = species,
                              y = change_in_density,
                              fill = geographic_range)) +
  geom_bar(stat = "identity") +
  scale_fill_paletteer_d("ghibli::MononokeMedium", direction = -1) +
  labs(x = "Species", y = "Change in Density", title = "Change in Species Density from 1955-2020", fill = "Geographic Range") + # fill changes legend title
  theme_minimal()

climate_geo_bar # PATTERN: Southern species increase, Northern species decreased.



# Visualize changes in density due to trophic level

climate_tro_bar <- ggplot(climate_change,
                          aes(x = species,
                              y = change_in_density,
                              fill = trophic_level)) +
  geom_bar(stat = "identity") +
  scale_fill_paletteer_d("ghibli::MononokeMedium", direction = -1) +
  theme_minimal()

climate_tro_bar

```

Species whose range exists south of a Northern California coastal site have increased over the past 65 years. On the other hand, species whose range is north of the study site have decreased. This pattern suggests that climatic changes have impacted marine communities in California. As climate change leads to global increases in sea surface temperature, species who are adapted to cooler water temperatures (those who have northern ranges) are experiencing decreases in density. The opposite happens to species who have southern ranges and are already adapted to warmer water temperatures.


## 2. Visualizing changes at nuclear power plant
```{r}

# Read in and wrangle data for nuclear power plant

nuclear_power <- read_csv("nuclear_power.csv") %>% 
  clean_names() %>% 
  gather(`hyperion_bay`,`noway_bay`,`diablo_cove`, key = "site", value = "density") # Density measured in (# of individuals/100m2)

```

```{r}

# Visualize change in all species across all sites

nuclear_species <- ggplot(nuclear_power, aes(x = year, y = density, fill = site)) +
  geom_line(aes(color = site)) +
  scale_color_paletteer_d("ghibli::MononokeMedium", direction = -1,
                          labels = c("Diablo Cove", "Hyperion Bay", "Noway Bay")) + # Change legend labels and keep exisitng palette
  facet_wrap(~species, scale = "free", nrow = 3) + # change number of rows for facet
  labs(x = "Year", y = expression(Mean~Density~("#"/100~m^2)), title = "Species Density Monitoring Program (2000-2019)", color = "Site") + # color changes legend title
  theme_minimal()

nuclear_species

```
El Diablo nuclear power plant seemd to slightly alter the density of species C in Diablo Bay. Shortly after the plant became operations, species C densities increased in Diablo Bay compared to Hyperion Bay. Hyperion Bay is used as a control in this analysis because it is the site most silimar to Diablo Bay before the plant became operational. In 2015, species C densities increased again in Hyperion Bay, and now the two Bays show similar patterns of density changes.


## 3. Visualizing patterns of change for crab fishery
```{r}

# Read in and wrangle data for crab fishery

crab_barnacle <- read_csv("crab_barnacle.csv") %>% 
  clean_names() %>% 
  gather(`cancer_ca`,`cancer_or`,`barnacle_ca`,`barnacle_or`, key = "site", value = "density") %>% 
  spread(stage, density) %>% 
  rename(adult_density = Adult) %>% # rename column
  rename(recruit_density = Recruit) %>% 
  separate(site, into = c("species", "site"), sep = "_") %>%  # separate species from site
  arrange(year, species)

```

```{r}

# Read in and wrangle data for upwelling indices

upwelling <- read_csv("upwelling.csv") %>% 
  clean_names() %>% 
  gather(`winter_ca`,`summer_ca`,`winter_or`,`summer_or`, key = "season", value = "upwelling_indices") %>% # gather seasons & put values in a separate col
  separate(season, into = c("season","site"), sep = "_") %>% 
  arrange(year)

```

```{r}

# Join crab_barnacle to upwelling

crab_fishery <- left_join(crab_barnacle, upwelling) %>% # join two dataframes, default wwill join by matching columns in this case, "year" and "site"
  clean_names() %>% 
  mutate(site = str_to_upper(site)) %>% # Make everything in site column uppercase
  arrange(year)

# crab_fishery$year <- as.character(crab_fishery$year) 

```

```{r}

# Subset of data for barnacles

barnacles <- crab_fishery %>% 
  filter(species == "barnacle")


# Subest of data for crabs

crabs <- crab_fishery %>% 
  filter(species == "cancer")

```

```{r, eval = FALSE}

# These comparisons were done in Excel instead. I could not figure out how to graph is such that values were offset by one year. (adult densities in one year affected recruit densities in the next year)


# Relationship betewwen density of adult and next year's recruits for barnacles

barnacles_plot <- ggplot(barnacles, aes(x = adult_density, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Adult Density", y = "Recruit Density", title = "Recruit v. Adult Densities in Barnacles", color = "Year") +
  theme_bw()

barnacles_plot



# Relationship between density of adult and next year's recruits in crabs

crabs_plot <- ggplot(crabs, aes(x = adult_density, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free") +
  theme_classic()

crabs_plot


```

```{r}

# Relationship between desnsity of recruits and winter upwelling indicies for barnacles

recruit_barnacles_upwelling <- barnacles %>% 
  filter(season == "winter")

barnacle_up_recruit_plot <- ggplot(recruit_barnacles_upwelling, aes(x = upwelling_indices, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Upwelling Index", y = "Recruit Density", title = "Recruit Density vs. Winter Upwelling Index for Barnacles", color = "Year") +
  theme_bw()

barnacle_up_recruit_plot



# Relationship between density of adults and upwelling indicies for barnacles

adult_barnacles_upwelling <- barnacles %>% 
  filter(season == "summer")

barnacle_up_adult_plot <- ggplot(adult_barnacles_upwelling, aes(x = upwelling_indices, y = adult_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Upwelling Index", y = "Adult Density", title = "Adult Density vs. Summer Upwelling Index for Barnacles", color = "Year") +
  theme_bw()

barnacle_up_adult_plot



# Relationship between density of recruits and upwelling indicies for crabs

recruit_crabs_upwelling <- crabs %>%
  filter(season == "winter")

crab_up_recruit_plot <- ggplot(recruit_crabs_upwelling, aes(x = upwelling_indices, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  theme_bw()

crab_up_recruit_plot



# Relationship between density of adults and upwelling indicies for crabs

adult_crabs_upwelling <- crabs %>% 
  filter(season == "summer")

crab_up_adult_plot <- ggplot(adult_crabs_upwelling, aes(x = upwelling_indices, y = adult_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Upwelling Index", y ="Adult Density", title ="Adult Density vs. Summer Upwelling Index for Crabs", color = "Year") +
  theme_bw()

crab_up_adult_plot

```
Winter upwelling indices affect recruit densities, because this is when adults are producing new eggs. On the other hand, summer upwelling indices affect adult densities, because this is when juveniles are developing.

There is a strong positive relationship between recruit density and the winter upwelling index for barnacles in California. Downwelling (negative values) decreases recruit density and upwelling (positive values) increases recruit density. This makes sense because downwelling is associated with low nutrients and rare phytoplankton. This, of course, would affect the adult population and therefore cause them to produce fewer eggs.









## 4. EXTRA

```{r}


# Grid plots to double check

barnacle_up_recruit_plot <- ggplot(barnacles, aes(x = upwelling_indices, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_grid(season~site, scale = "free") +
  labs(x = "Upwelling Index", y = "Recruit Density", title = "Recruit Density vs. Upwelling Index for Barnacles", color = "Year") +
  theme_bw()

barnacle_up_recruit_plot




barnacle_up_adultt_plot <- ggplot(barnacles, aes(x = upwelling_indices, y = adult_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_grid(season~site, scale = "free") +
  labs(x = "Upwelling Index", y = "Adult Density", title = "Adult Density vs. Upwelling Index for Barnacles", color = "Year") +
  theme_bw()

barnacle_up_adultt_plot






crab_up_recruit_plot <- ggplot(crabs, aes(x = upwelling_indices, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_grid(season~site, scale = "free") +
  theme_bw()

crab_up_recruit_plot



crab_up_adultt_plot <- ggplot(crabs, aes(x = upwelling_indices, y = adult_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_grid(season~site, scale = "free") +
  theme_bw()

crab_up_adultt_plot




```

