---
title: "HW4"
author: "Corinna Hong"
date: "February 24, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center")
```

```{r}

# load packages

library(tidyverse)
library(janitor)
library(paletteer)

```

##1. Visualizing patterns of change for climate change data
```{r}

# read in and wrangle climate change data

climate_change <- read_csv("climate_change.csv") %>% 
  clean_names() %>% 
  mutate(geographic_range = str_to_lower(geographic_range))

```

```{r}

# Visualize changes in density due to development mode

climate_dev_bar <- ggplot(climate_change,
                          aes(x = species,
                              y = change_in_density,
                              fill = development_mode)) +
  geom_bar (stat = "identity") + # identity means height of bars equals actual value
   scale_fill_paletteer_d("ghibli::MononokeMedium", direction = -1) + # direction -1 means start from the right side of the palette
  theme_minimal()

climate_dev_bar



# Visualize changes in density due to geographic range

climate_geo_bar <- ggplot(climate_change,
                          aes(x = species,
                              y = change_in_density,
                              fill = geographic_range)) +
  geom_bar(stat = "identity") +
  scale_fill_paletteer_d("ghibli::MononokeMedium", direction = -1) +
  labs(x = "Species", y = "Change in Density", title = "Change in Species Density from 1955-2020", fill = "Geographic Range") + # fill changes legend title
  theme_minimal()

climate_geo_bar # PATTERN: Southern species increase, Northern species decreased.



# Visualize changes in density due to trophic level

climate_tro_bar <- ggplot(climate_change,
                          aes(x = species,
                              y = change_in_density,
                              fill = trophic_level)) +
  geom_bar(stat = "identity") +
  scale_fill_paletteer_d("ghibli::MononokeMedium", direction = -1) +
  theme_minimal()

climate_tro_bar

```

Species whose range exists south of a Northern California coastal site have increased over the past 65 years. On the other hand, species whose range is north of the study site have decreased. This pattern suggests that climatic changes have impacted marine communities in California. As climate change leads to global increases in sea surface temperature, species who are adapted to cooler water temperatures (those who have northern ranges) are experiencing decreases in density. The opposite happens to species who have southern ranges and are already adapted to warmer water temperatures. 


## 2. Visualizing changes at nuclear power plant
```{r}

# Read in and wrangle data for nuclear power plant

nuclear_power <- read_csv("nuclear_power.csv") %>% 
  clean_names() %>% 
  gather(`hyperion_bay`,`noway_bay`,`diablo_cove`, key = "site", value = "density") # Density measured in (# of individuals/100m2)

```

```{r}

# Visualize change in all species across all sites

nuclear_species <- ggplot(nuclear_power, aes(x = year, y = density, fill = site)) +
  geom_line(aes(color = site)) +
  scale_color_paletteer_d("ghibli::MononokeMedium", direction = -1,
                          labels = c("Diablo Cove", "Hyperion Bay", "Noway Bay")) + # Change legend labels and keep exisitng palette
  facet_wrap(~species, scale = "free", nrow = 3) + # change number of rows for facet
  labs(x = "Year", y = expression(Mean~Density~("#"/100~m^2)), title = "Species Density Monitoring Program (2000-2019)", color = "Site") + # color changes legend title
  theme_minimal()

nuclear_species

```
El Diablo nuclear power plant seemd to slightly alter the density of species C in Diablo Bay. Shortly after the plant became operations, species C densities increased in Diablo Bay compared to Hyperion Bay. Hyperion Bay is used as a control in this analysis because it is the site most silimar to Diablo Bay before the plant became operational. In 2015, species C densities increased again in Hyperion Bay, and now the two Bays show similar patterns of density changes. 


```{r}

# Read in and wrangle data for crab fishery

crab_barnacle <- read_csv("crab_barnacle.csv") %>% 
  clean_names() %>% 
  gather(`cancer_ca`,`cancer_or`,`barnacle_ca`,`barnacle_or`, key = "site", value = "density") %>% 
  spread(stage, density) %>% 
  rename(adult_density = Adult) %>% # rename column
  rename(recruit_density = Recruit) %>% 
  separate(site, into = c("species", "site"), sep = "_") %>%  # separate species from site
  arrange(year, species)

```

```{r}

# Read in and wrangle data for upwelling indices

upwelling <- read_csv("upwelling.csv") %>% 
  clean_names() %>% 
  gather(`winter_ca`,`summer_ca`,`winter_or`,`summer_or`, key = "season", value = "upwelling_indices") %>% # gather seasons & put values in a separate col
  separate(season, into = c("season","site"), sep = "_") %>% 
  arrange(year)

```

```{r}

# Join crab_barnacle to upwelling

crab_fishery <- left_join(crab_barnacle, upwelling) %>% # join two dataframes, default wwill join by matching columns in this case, "year" and "site"
  clean_names() %>% 
  mutate(site = str_to_upper(site)) %>% # Make everything in site column uppercase
  arrange(year)

# crab_fishery$year <- as.character(crab_fishery$year) 

```

## 3. Visualizing patterns of change for crab fishery
```{r}

# Relationship betewwen density of adult and next year's recruits for barnacles

barnacles <- crab_fishery %>% 
  filter(species == "barnacle")

barnacles_plot <- ggplot(barnacles, aes(x = adult_density, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Adult Density", y = "Recruit Density", title = "Recruit v. Adult Densities in Barnacles", color = "Year") +
  theme_bw()

barnacles_plot



# Relationship between density of adult and next year's recruits in crabs

crabs <- crab_fishery %>% 
  filter(species == "cancer")

crabs_plot <- ggplot(crabs, aes(x = adult_density, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free") +
  theme_classic()

crabs_plot


```
 
 Slight relationship between recruit and adult densities in CA for barnacles.


```{r}

# Relationship between desnsity of recruits and upwelling indicies for barnacles

recruit_barnacles_upwelling <- barnacles %>% 
  filter(season == "summer")

barnacle_up_recruit_plot <- ggplot(recruit_barnacles_upwelling, aes(x = upwelling_indices, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Upwelling Index", y = "Recruit Density", title = "Recruit Density vs. Summer Upwelling Index for Barnacles", color = "Year") +
  theme_bw()

barnacle_up_recruit_plot



# Relationship between density of adults and upwelling indicies for barnacles

adult_barnacles_upwelling <- barnacles %>% 
  filter(season == "winter")

barnacle_up_adult_plot <- ggplot(adult_barnacles_upwelling, aes(x = upwelling_indices, y = adult_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Upwelling Index", y = "Adult Density", title = "Adult Density vs. Winter Upwelling Index for Barnacles", color = "Year") +
  theme_bw()

barnacle_up_adult_plot



# Relationship between density of recruits and upwelling indicies for crabs

recruit_crabs_upwelling <- crabs %>%
  filter(season == "summer")

crab_up_recruit_plot <- ggplot(recruit_crabs_upwelling, aes(x = upwelling_indices, y = recruit_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free") +
  theme_bw()

crab_up_recruit_plot



# Relationship between density of adults and upwelling indicies for crabs

adult_crabs_upwelling <- crabs %>% 
  filter(season == "winter")

crab_up_adult_plot <- ggplot(adult_crabs_upwelling, aes(x = upwelling_indices, y = adult_density)) +
  geom_point(aes(color = year)) +
  scale_color_paletteer_c("gameofthrones::lannister", direction = -1) +
  facet_wrap(~site, scale = "free", nrow = 2) +
  labs(x = "Upwelling Index", y ="Adult Density", title ="Adult Density vs. Winter Upwelling Index for Crabs", color = "Year") +
  theme_bw()

crab_up_adult_plot

```

Slight relationship between recruit density and summer upwelling index for barnacles in cA
Slight relationship between adult density and winter upwelling index for barnacles in cA
Slight relationship between adult density and winter upwelling index for crabs in CA ?
Slight relaiontship between adult density and winter upwelling index for crabs in OR?